Report and analysis on implementation of IMM algorithm for multiple-model dynamics tracking

### 2020-02-02 Latex/draft.tex
Onlyimages missing

    \item ON: at each time-step checks if it is still in range with the function inRange(), takes a measurement, computes the IMM algorithm and
          at a tunable rate computes the consensus with the other nearby ON sensors. If it's not in range anymore then turns IDLE.
    \item IDLE : checks if it is now in range with the function inRange(), if it is then initialize itself with the data from nearby sensors

@@

The model choosen for our sensor is a Radar, and as such the non-linear measurement function

@@

cordinates relative to the sensor), $z(k)$ is the measure and $v(k)$ is the noise associated to the measurement's operation. The measurement noise $w(k)$ is associated

@@

\section*{Model used}
As mentioned already, the agent we want to track has a variable-model dynamic that can switch between different elements in a set
of models that we denote as $\mathbbm{M}$. We will consider 2 families of such sets that we will call: Random Accelerated Walk $\mathbbm{M}_{1}$
and Random Unicycle Turning $\mathbbm{M}_{2}$. Every element of said families is a discrete Markovian process, as all are influenced by a white
noise on their inputs.
\subsection*{Random Accelerated Walk}
For the Random Accelerated Walk we consider the set $\mathbbm{M}_{1}$ to be 5 elements large.
Here the elements are described
\begin{center}
    \begin{tabular}{||c||c |c |}%{|width=\columnwidth|}
        \hline
        Random Walk                         \\
        \hline\hline
        Mode   & Behaviour                  \\ [0.5ex]
        \hline\hline
        Mode 1 & Constant Speed             \\
        \hline
        Mode 2 & Positive Acceleration in x \\
        \hline
        Mode 3 & Negative Acceleration in x \\
        \hline
        Mode 4 & Positive Acceleration in y \\
        \hline
        Mode 5 & Negative Acceleration in y \\ [1ex]
        \hline
    \end{tabular}
\end{center}

@@

\begin{equation}

@@

\end{equation}

@@

tracking and we have to rely on the Interacting Multiple Models algorithm (IMM). In essence the algorithm let various estimation filters work in parallel
(EKF in our case), and outputs a mixture of the filters' estimate based on the probability of each of the models used in the filter to be correct.

@@

Following is a short recapof the algorithm, to see it in much more detail refer to %\cite%

@@

           In this step a lower bound on the likelihood is implemented since a likelihood vector of $0$ breaks the algorithm.

@@

          the outputs. This last step is only for the sake of the output, the algorithm forms a loop without it.

@@

@@

\newpage

@@

Here we present the results of our implementation of the algorithm and sensors. The agent was left wandering inside the set virtual room 
and measured at each timestep the estimation error that the algorithm gave. The room is set to be a square with a $100m$ side with a $10\times 10$ 
grid of sensors with range$=10$. 

@@

different $R$s and $Q$s and we considered both the errors given by the Consensus and by each individual sensor's IMM. 

@@

and while $Q2=diag(1,1)$ is shared by $\mathbbm{M}_{1}$ and $\mathbbm{M}_{2}$, $Q1=diag(0.2,0.2)$ for $\mathbbm{M}_{2}$ and $Q1=diag(0.1,0.1)$ for $\mathbbm{M}_{1}$.

@@

What we gather from these tables is that, specifically in the case of the unicycle $\mathbbm{M}_{2}$ (that is a more complex model set), 
a consensus computed at a high frequency (rate=1, rate=2) gives disastrous results with any magnitude of noise. This is probably caused by 
the nature of the IMM algorithm that gets impeded by the consensus which homogenizes the internal Kalman filters input to a point which 
the IMM can't compute their probabilities efficiently. In fact we see a much better results with lower frequency such as rate=10 and rate=20.
We also tried the case in which we do not compute any consensus and we let individual sensors work alone and, while we see good results, 
it doesn't seem to be the best compromise.
\\
In the case of the Random Walk with a rate=2 we have good results with low $R$, but the algorithm is really sensible to changes to this 
parameter and performs really bad with $R1$.
\\
While we see that with both model sets the case with rate=20 gives very good results, but we also notice that the maximum of both the individual 
IMMs and of the consensus can get pretty high so if we have an application where the frequency of estimation must be high and we can't 
have a big overshoot, then this rate can fail.
All things considered the best compromise may be the rate=10, which always gives good results and has double the frequency of rate=20.
\\
Following are some graphs of the best and worst cases computed by the simulations. More graphs and a video of the algorithm working can 
be found at ...

>>>>

    \item ON: at each time-step checks if the agents is still in range with the function inRange(), takes a measurement, computes the IMM algorithm and
          at a tunable rate computes the consensus with the other nearby ON sensors. If the agent is not in range anymore then turns IDLE.
    \item IDLE : checks if the agent is now in range with the function inRange(), if it is then initialize itself with the data from nearby sensors

@@

The model chosen for our sensor is that of a radar, and as such the non-linear measurement function

@@

cordinates relative to the sensor), $z(k)$ is the measure and $v(k)$ is the noise associated to the measurement's operation. The measurement noise $v(k)$ is associated

@@

\section*{Model used}
As mentioned already, the agent we want to track has a variable-model dynamic that can switch between different elements in a set
of models that we denote as $\mathbbm{M}$. We will consider 2 families of such sets that we will call: Random Accelerated Walk $\mathbbm{M}_{1}$
and Random Unicycle Turning $\mathbbm{M}_{2}$. Every element of said families is a discrete Markovian process, as all are influenced by
noise on their inputs.
\subsection*{Random Accelerated Walk}
For the Random Accelerated Walk we consider the set $\mathbbm{M}_{1}$ to be 5 elements large.
Here the elements are described
        Random Walk                         \\
        Mode 1 & Constant Speed             \\
        Mode 2 & Positive Acceleration in x \\
        Mode 3 & Negative Acceleration in x \\
        Mode 4 & Positive Acceleration in y \\
        Mode 5 & Negative Acceleration in y \\ [1ex]

@@

@@

@@

tracking and we have to rely on the Interacting Multiple Models algorithm (IMM). In essence the algorithm let various estimation filters work in parallel, and outputs a mixture of the filters' estimate based on the probability of each of the models used in the filter to be correct.

@@

Following is a short recap of the algorithm, to see it in much more detail refer to CITA QUA

@@

           In this step two adjustment are needed: the first one is the choice to use a logaritmic pdf instead of a normal one due 
           to the fact that it gives better numerical results and lower bound on the likelihood is implemented since a likelihood 
           vector of $0$ breaks the algorithm.

@@

          the outputs. This last step is only for the sake of the output, the algorithm forms a loop without it as the starting step of 
          the next iteration uses the outputs of the filters at the steps before this one.

@@

The same WLS algorithm is used to initialize IDLE sensors that turn ON using data from nearby ON sensors.

@@

@@

Here we present the results of our implementation of the algorithm and sensors' grid. The agent was left wandering inside the set virtual room 
and at each timestep the estimation error of the algorithm was measured. The room is set to be a square with a $100m$ side with a $10\times 10$ 
grid of sensors with range$=10m$. 

@@

different $R$ and $Q$ and we considered both the errors given by the Consensus and by each individual sensor's IMM. We've also run simulations 
with noconsensus at all.

@@

and while $Q2=diag(1,1)$ is shared by $\mathbbm{M}_{1}$ and $\mathbbm{M}_{2}$, $Q1=diag(0.2,0.2)$ for $\mathbbm{M}_{2}$ and $Q1=diag(0.1,0.1)$ for 
$\mathbbm{M}_{1}$. The Transition matrices were initializated as follows
    &T_{1}=\begin{bmatrix}
        0.6&0.1&0.1&0.1&0.1\\
        0.2&0.5&0.12&0.06&0.06\\
        0.2&0.12&0.5&0.12&0.06\\
        0.2&0.06&0.12&0.5&0.12\\
        0.2&0.12&0.06&0.12&0.5\\
    \end{bmatrix}\\
    &T_{2}=\begin{bmatrix}
        0.8&0.05&0.05&0.05&0.05\\
        0.25&0.5&0.25&0&0\\
        0.25&0.25&0.5&0&0\\
        0.25&0&0&0.5&0.25\\
        0.25&0&0&0.25&0.5\\

@@

What we gather from these tables is that, specifically in the case of the unicycle $\mathbbm{M}_{2}$ (that is a more complex model set), 
a consensus computed at a high frequency (rate=1, rate=2) gives disastrous results with any magnitude of noise. This is probably caused by 
the nature of the IMM algorithm that gets impeded by the consensus which homogenizes the internal Kalman filters input to a point which 
the IMM can't compute their probabilities efficiently. In fact we see a much better results with lower frequency such as rate=10 and rate=20.
We also tried the case in which we do not compute any consensus and we let individual sensors work alone and, while we see good results, 
it doesn't seem to be the best compromise.
\\
In the case of the Random Walk with a rate=2 we have good results with low $R$, but the algorithm is really sensible to changes to this 
parameter and performs really bad with $R1$.
\\
While we see that with both model sets the case with rate=20 gives very good results, but we also notice that the maximum of both the individual 
IMMs and of the consensus can get pretty high so if we have an application where the frequency of estimation must be high and we can't 
have a big overshoot, then this rate can fail.
All things considered the best trade-off may be the rate=10, which always gives good results and has double the frequency of rate=20.
\\
Following are some graphs of the best and worst cases computed by the simulations. More graphs and a video of the algorithm working can 
be found at ...
